---
title: "Hosting data for JBrowseR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Hosting data for JBrowseR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

Every JBrowseR function that accepts data takes a URL. Genomics files are far
too large to inline as Shiny static resources, and the browser fetches only the
slices it needs directly from the host — so the data lives at a URL and JBrowse
reads from it in place.

For human and model organisms you often need no hosting at all: pass a hub name
straight to `JBrowseR()` and the assembly, reference-name aliases, cytobands, and
gene-name search all load from the CORS-enabled hub at
[jbrowse.org](https://jbrowse.org).

```{r}
JBrowseR("hg38", location = "BRCA1")
```

For your own data you host the files yourself. On the JBrowse 2 team we use
[Amazon S3](https://aws.amazon.com/s3/) — each object has its own URL — but any
web server works, e.g. [Apache](https://httpd.apache.org/) or a cloud bucket.

## Server requirements

Whichever host you pick, it must satisfy two requirements:

- **HTTP `Range` requests**, so the browser can use
  [byte serving](https://en.wikipedia.org/wiki/Byte_serving) to fetch only the
  chunks needed to render the current view instead of the whole file.
- **CORS**, because the request originates from wherever the browser is running,
  not from where the data is hosted.

Here is an example CORS configuration for an S3 bucket:

```{js}
[
    {
        "AllowedHeaders": [
            "*"
        ],
        "AllowedMethods": [
            "GET"
        ],
        "AllowedOrigins": [
            "*"
        ],
        "ExposeHeaders": [
            "Accept-Ranges",
            "Content-Range",
            "Content-Encoding",
            "Content-Length"
        ],
        "MaxAgeSeconds": 3000
    }
]
```

If you use S3 and plan to deploy with [shinyapps.io](https://www.shinyapps.io/),
put the bucket in Amazon's US-East region — shinyapps.io is hosted entirely
there, so co-locating the data gives faster performance.

## Viewing local files

Files on your own machine need no host at all. `local_files` reads them into the
document, and a track then refers to one by name exactly as it would a URL:

```{r}
JBrowseR(
  "hg38",
  tracks = list(list(uri = "peaks.bed.gz", name = "Peaks")),
  local_files = "~/data/peaks.bed.gz",
  location = "chr1:1-10000"
)
```

The name is the file's basename, so the `.bed.gz` extension still picks the
adapter. A conventional sibling index next to the file — `.tbi`, `.csi`, `.bai`,
`.crai`, `.fai`, `.gzi` — is registered alongside it under the name the adapter
derives, so an indexed file stays indexed: the browser seeks into the bytes it
holds and reads only the region on screen, the same as it would over HTTP.

This is the one case where no server is involved, so the two requirements above
do not apply — there is no origin to configure CORS for, and the range reads
happen inside the browser. It is also what makes a knitted document
self-contained: the bytes travel with the HTML, where a `http://localhost:...`
URL would be dead for anyone you send it to.

The cost is that the bytes ride inside the document, base64-encoded, rather than
being fetched a slice at a time. That is fine for a file on an analyst's laptop
and wrong for a reference genome — `local_files` warns past ~50 MB. Host
anything larger and pass its URL.

An assembly can come from a local file the same way, since `uri` is a name here
like any other:

```{r}
JBrowseR(
  assembly = list(name = "mine", uri = "genome.fa.gz"),
  local_files = c("~/data/genome.fa.gz", "~/data/genes.gff.gz"),
  tracks = list(list(uri = "genes.gff.gz", name = "Genes"))
)
```
