---
title: "Hosting data for JBrowseR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Hosting data for JBrowseR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

Every JBrowseR function that accepts data takes a URL. Genomics files are far
too large to inline as Shiny static resources, and the browser fetches only the
slices it needs directly from the host — so the data lives at a URL and JBrowse
reads from it in place.

For human and model organisms you often need no hosting at all: pass a hub name
straight to `JBrowseR()` and the assembly, reference-name aliases, cytobands, and
gene-name search all load from the CORS-enabled hub at
[jbrowse.org](https://jbrowse.org).

```{r}
JBrowseR("hg38", location = "BRCA1")
```

For your own data you host the files yourself. On the JBrowse 2 team we use
[Amazon S3](https://aws.amazon.com/s3/) — each object has its own URL — but any
web server works, e.g. [Apache](https://httpd.apache.org/) or a cloud bucket.

## Server requirements

Whichever host you pick, it must satisfy two requirements:

- **HTTP `Range` requests**, so the browser can use
  [byte serving](https://en.wikipedia.org/wiki/Byte_serving) to fetch only the
  chunks needed to render the current view instead of the whole file.
- **CORS**, because the request originates from wherever the browser is running,
  not from where the data is hosted.

Here is an example CORS configuration for an S3 bucket:

```{js}
[
    {
        "AllowedHeaders": [
            "*"
        ],
        "AllowedMethods": [
            "GET"
        ],
        "AllowedOrigins": [
            "*"
        ],
        "ExposeHeaders": [
            "Accept-Ranges",
            "Content-Range",
            "Content-Encoding",
            "Content-Length"
        ],
        "MaxAgeSeconds": 3000
    }
]
```

If you use S3 and plan to deploy with [shinyapps.io](https://www.shinyapps.io/),
put the bucket in Amazon's US-East region — shinyapps.io is hosted entirely
there, so co-locating the data gives faster performance.

## Viewing local files

To view files from your own machine, serve the directory with any static server
that supports range requests and CORS, then point `assembly()` / `track()` at
the resulting `http://localhost:...` URLs. For example, with
[http-server](https://www.npmjs.com/package/http-server):

```sh
npx http-server --cors -p 5000 ~/path/to/my-data
```

```{r}
JBrowseR(
  assembly = assembly("http://127.0.0.1:5000/genome.fa.gz"),
  tracks = tracks(track("http://127.0.0.1:5000/genes.gff.gz", name = "Genes")),
  location = "chr1:1-10000"
)
```
