---
title: "Custom browser tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Custom browser tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

When your genome is one of the hosted hubs, `JBrowseR("hg38")` is all you need.
This tutorial covers the other case: building a browser for a genome you host
yourself, with your own tracks, gene-name search, and theme.

```{r setup}
library(JBrowseR)
```

## Describe the assembly

An assembly is a list with a `name` and a `uri`. JBrowse derives the index
locations (`.fai`, plus `.gzi` for bgzipped FASTA) from the URL, so you only
point at the FASTA itself. Add reference-name aliases so `chr1`/`1` both resolve.

```{r}
hg19 <- list(
  name = "hg19",
  uri = "https://jbrowse.org/genomes/hg19/fasta/hg19.fa.gz",
  aliases = list("GRCh37"),
  refNameAliases = list(uri = "https://jbrowse.org/genomes/hg19/hg19_aliases.txt")
)
```

## Add tracks

A track is a list too. `uri` alone is enough — the view infers the track type
and adapter from the file extension, and derives the index location. You do not
need to set `assemblyNames` on each track; the view fills it in from the
assembly you load.

```{r}
my_tracks <- list(
  list(
    uri =
    "https://jbrowse.org/genomes/hg19/GRCh37_latest_genomic.sort.gff.gz",
    name = "NCBI RefSeq Genes"
  )
)
```

## Enable gene-name search

Hub assemblies include search; for a custom assembly, point at your own Trix
index files with a Trix adapter and pass it as `text_search`. Now `location` can
be a gene name.

```{r}
hg19_search <- list(
  type = "TrixTextSearchAdapter",
  textSearchAdapterId = "hg19-index",
  assemblyNames = list("hg19"),
  ixFilePath = list(uri = "https://jbrowse.org/genomes/hg19/trix/hg19.ix"),
  ixxFilePath = list(uri = "https://jbrowse.org/genomes/hg19/trix/hg19.ixx"),
  metaFilePath = list(uri = "https://jbrowse.org/genomes/hg19/trix/meta.json")
)
```

## Put it together

```{r}
JBrowseR(
  assembly = hg19,
  tracks = my_tracks,
  text_search = hg19_search,
  theme = list(palette = list(primary = list(main = "#311b92"))),
  location = "MYC"
)
```

## Show results computed in R

`track_data_frame()` turns a data frame into an in-browser track with no files
and no server — the natural way to put an analysis you ran in R onto the genome.
The frame needs `chrom`, `start`, `end`, and `name` columns; an optional `score`
column makes it a quantitative track.

```{r}
regions <- data.frame(
  chrom = c("10", "10"),
  start = c(29838737, 29850000),
  end   = c(29840000, 29855000),
  name  = c("regionA", "regionB"),
  score = c(42, 88)
)

JBrowseR(
  assembly = hg19,
  tracks = list(track_data_frame(regions, "my_regions")),
  location = "10:29,838,737..29,855,000"
)
```

## Reacting to clicks in Shiny

When rendered inside Shiny, clicking a feature sets `input$selectedFeature` to
the feature's data, so you can build tables, plots, or links from the current
selection.

```{r}
# server side
output$browser <- renderJBrowseR(
  JBrowseR(assembly = hg19, tracks = my_tracks, location = "MYC")
)
observeEvent(input$selectedFeature, {
  print(input$selectedFeature$name)
})
```
