---
title: "Advanced JSON tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Advanced JSON tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

The [introduction](JBrowseR.html) and
[custom browser tutorial](custom-browser-tutorial.html) cover the declarative
API, which is enough for most browsers. This tutorial covers the escape hatch:
handing `JBrowseR()` a whole JBrowse 2 configuration through the `config`
argument, for full control over every option.

```{r setup}
library(JBrowseR)
```

## When to use it

Reach for `config` when you need something the helpers don't expose — a saved
`defaultSession`, per-track display options, internet accounts for authenticated
data, or runtime plugins. A JBrowse `config` mirrors the format documented in
the [JBrowse config guide](https://jbrowse.org/jb2/docs/config_guide) and may
contain `assembly`, `tracks`, `defaultSession`, `aggregateTextSearchAdapters`,
`theme`, and `plugins`.

## From a config file

`config` takes the path or URL of a `config.json` and reads it for you (JSON
text works too). A config file is portable — the same file works in the JBrowse
web app and desktop.

```{r}
JBrowseR(config = "https://jbrowse.org/genomes/hg19/config.json")
```

## From an R list

Because htmlwidgets serializes R lists to JSON, you can also assemble the config
inline. Mix in the declarative helpers wherever they are convenient — they
return plain lists.

```{r}
cfg <- list(
  assembly = assembly(
    "https://jbrowse.org/genomes/hg19/fasta/hg19.fa.gz",
    aliases = "GRCh37"
  ),
  tracks = tracks(
    track(
      "https://jbrowse.org/genomes/hg19/GRCh37_latest_genomic.sort.gff.gz",
      name = "NCBI RefSeq Genes"
    )
  ),
  defaultSession = list(
    name = "My session",
    view = list(
      type = "LinearGenomeView",
      tracks = list(list(
        type = "FeatureTrack",
        configuration = "NCBI RefSeq Genes"
      ))
    )
  )
)

JBrowseR(config = cfg, location = "10:29,838,737..29,838,819")
```

`location` still applies alongside a `config`, unless the config's
`defaultSession` already positions the view.
