---
title: "Advanced JSON tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Advanced JSON tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

The [introduction](JBrowseR.html) and
[custom browser tutorial](custom-browser-tutorial.html) pass JBrowse config to
`JBrowseR()` one argument at a time. The `config` argument takes the same
options as one JSON object instead, so a browser can live in a file that R,
Python and JavaScript all read.

```{r setup}
library(JBrowseR)
```

## What a config holds

`config` is `JBrowseR()`'s options in the camelCase JBrowse spells them:
`assembly`, `tracks`, `location`, `session`, `aggregateTextSearchAdapters`,
`configuration`, `plugins`, and `internetAccounts`, the one with no R argument
of its own, for authenticated data. Arguments passed beside `config` override
the fields it holds.

It is not a JBrowse Web `config.json`. That file lists `assemblies` in the
plural, opens a `defaultSession`, and names data relative to its own URL, and
`JBrowseR()` reads none of the three. To open a genome JBrowse hosts, name its
hub instead — `JBrowseR("hg19")` fetches `https://jbrowse.org/ucsc/hg19/config.json`
and resolves its paths.

## From a config file

`config` takes the path or URL of a JSON file and reads it for you (JSON text
works too). The
[`load_config_json`](https://github.com/GMOD/JBrowseR/tree/main/example_apps/load_config_json)
example app keeps its assembly in one:

```{r}
JBrowseR(config = "config.json", location = "10:29,838,737..29,838,819")
```

## From an R list

Because htmlwidgets serializes R lists to JSON, the same object can be written
inline:

```{r}
cfg <- list(
  assembly = list(
    name = "hg19",
    uri = "https://jbrowse.org/genomes/hg19/fasta/hg19.fa.gz",
    aliases = list("GRCh37")
  ),
  tracks = list(
    list(
      uri = "https://jbrowse.org/genomes/hg19/GRCh37_latest_genomic.sort.gff.gz",
      name = "NCBI RefSeq Genes"
    )
  ),
  location = "10:29,838,737..29,838,819"
)

JBrowseR(config = cfg)
```
