---
title: "Advanced JSON tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Advanced JSON tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

`JBrowseR()` takes JBrowse's `createLinearGenomeView` options as named
arguments, and `JBrowseRApp()` takes `createApp`'s, both under the camelCase
names JBrowse gives them. A browser can therefore live in a JSON file that R,
Python and JavaScript all read.

```{r setup}
library(JBrowseR)
```

## What the options hold

`JBrowseR()`'s options are `assembly`, `tracks`, `location`, `session`,
`aggregateTextSearchAdapters`, `internetAccounts`, `plugins` and
`configuration`. The package passes them through unread, so an option JBrowse
adds works here without a package update.

A JBrowse Web `config.json` is not that shape: it lists `assemblies` in the
plural and opens a `defaultSession`. Its `assemblies`, `tracks` and `plugins`
are `JBrowseRApp()` options, though, and its `defaultSession.views` is
`JBrowseRApp()`'s `views`. To open a genome JBrowse hosts, name its hub instead —
`JBrowseR(assembly = "hg19")` fetches `https://jbrowse.org/ucsc/hg19/config.json`
and resolves its paths.

## From a JSON file

`do.call()` hands a list's entries to a function as named arguments, so a file
read with `jsonlite::read_json()` is a whole browser. The
[`load_config_json`](https://github.com/GMOD/JBrowseR/tree/main/example_apps/load_config_json)
example app keeps its assembly in one:

```{r}
do.call(JBrowseR, jsonlite::read_json("config.json"))
```

`modifyList()` overrides a field the file holds:

```{r}
opts <- jsonlite::read_json("config.json")
do.call(JBrowseR, modifyList(opts, list(location = "10:29,838,737..29,838,819")))
```

`read_json()` keeps JSON arrays as lists, so a one-element `assemblyNames`
stays an array on the way back out.

## From an R list

The same object written inline:

```{r}
opts <- list(
  assembly = list(
    name = "hg19",
    uri = "https://jbrowse.org/genomes/hg19/fasta/hg19.fa.gz",
    aliases = list("GRCh37")
  ),
  tracks = list(
    list(
      uri = "https://jbrowse.org/genomes/hg19/GRCh37_latest_genomic.sort.gff.gz",
      name = "NCBI RefSeq Genes"
    )
  ),
  location = "10:29,838,737..29,838,819"
)

do.call(JBrowseR, opts)
```
