---
title: "Live embedded browsers"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Every browser on this page is a **live, interactive JBrowse 2** — pan, zoom, and
click features directly in the page. This article is part of the website only
(it is not a package vignette), so it can embed the full widget; the
[Introduction](../JBrowseR.html) uses screenshots to keep the package small.

```{r setup}
library(JBrowseR)
```

## A hosted genome

Name a genome hub and you get the assembly, reference-name aliases, cytobands,
and gene-name search for free. Here `location = "BRCA1"` resolves through the
hub's search index.

```{r hub}
JBrowseR(
  assembly = "hg38",
  tracks = list(list(
    uri =
    paste0(
      "https://jbrowse.org/genomes/GRCh38/ncbi_refseq/",
      "GCA_000001405.15_GRCh38_full_analysis_set.refseq_annotation.sorted.gff.gz"
    ),
    name = "NCBI RefSeq Genes"
  )),
  location = "BRCA1"
)
```

## A custom genome from a FASTA URL

No hub needed for your own genome: pass a sequence-file URL straight to
`assembly`. The view builds the assembly from it (naming it after the file), and
a bare data-file URL in `tracks` is inferred into a track — here an alignments
track from a BAM.

```{r custom}
JBrowseR(
  assembly = "https://jbrowse.org/genomes/volvox/volvox.fa",
  tracks = list("https://jbrowse.org/genomes/volvox/volvox.bam"),
  location = "ctgA:1..5,000"
)
```

## An R result as a track

Turn a data frame of features into a track with `track_data_frame()` — no files,
no web server. A `score` column makes it quantitative.

```{r dataframe}
peaks <- data.frame(
  chrom = "17",
  start = seq(43000000, 43120000, by = 12000),
  end = seq(43000000, 43120000, by = 12000) + 4000,
  name = paste0("peak", seq_len(11)),
  score = round(runif(11, 5, 100))
)

JBrowseR(
  assembly = "hg38",
  tracks = list(track_data_frame(peaks, "R_peaks")),
  location = "17:43,000,000..43,125,000"
)
```

## A plot from a data frame

A `displays` entry says how to draw the rows. `LinearMarkDisplay` is JBrowse's
grammar of graphics: `encoding` maps columns to channels the way `aes()` does,
and the axis, legend and reference lines come from the scales. A windowed Fst
scan between two *Drosophila* populations, computed in base R from hosted allele
frequencies, becomes a bar per window on a value axis, coloured by a threshold
scale over the same column; the sweep at *Cyp6g1* is the run of bars past the
0.25 rule.

```{r marks}
freqs <- read.csv("https://jbrowse.org/demos/popgen/dest_cyp6g1_freqs.csv")
p1 <- freqs$afr_freq
p2 <- freqs$cosmo_freq
window <- freqs$pos %/% 10000 * 10000
num <- tapply((p1 - p2)^2, window, sum)
den <- tapply(p1 * (1 - p2) + p2 * (1 - p1), window, sum)
fst <- data.frame(
  chrom = "chr2R",
  start = as.integer(names(num)),
  end = as.integer(names(num)) + 10000,
  fst = round(pmax(num / den, 0), 3),
  n_snps = as.vector(table(window))
)
fst <- fst[fst$n_snps >= 20, ]

JBrowseR(
  assembly = "dm6",
  tracks = list(track_data_frame(
    fst, "Fst (African vs cosmopolitan)",
    displays = list(list(
      type = "LinearMarkDisplay",
      scales = list(y = list(title = "Fst", rules = list(0.25))),
      marks = list(list(
        shape = "bar",
        encoding = list(
          y = "fst",
          color = list(
            field = "fst", scale = "threshold",
            domain = list(0.12, 0.25),
            range = list("#90a4ae", "#f9a825", "#d84315")
          )
        )
      ))
    ))
  )),
  location = "chr2R:11,900,000..12,450,000"
)
```

## A file from this R session, with no server

`track_data_frame()` above puts every feature in the page as JSON, which stops
scaling somewhere past a few thousand of them. `local_files` is the other way to
have no web server: write a real file and hand JBrowse its bytes. Each path
registers under its basename, and that name is what a track's `uri` refers to —
so the `.bed` extension still picks the adapter, exactly as a URL's would.

```{r localfile}
bed <- file.path(tempdir(), "peaks.bed")
write.table(
  peaks[c("chrom", "start", "end", "name")],
  bed,
  sep = "\t", quote = FALSE, row.names = FALSE, col.names = FALSE
)

JBrowseR(
  assembly = "hg38",
  tracks = list(list(uri = "peaks.bed", name = "Peaks from a file")),
  local_files = bed,
  location = "17:43,000,000..43,125,000"
)
```

Nothing is served here: the bytes travel inside this page and JBrowse reads them
out of a `Blob`. That also means an **indexed** file stays indexed, which is the
reason to reach for this rather than inline the features — a bgzipped+tabixed
table, or a BAM with its `.bai`, is read by byte range, so only the region on
screen is ever touched. Put the index next to the data file and it comes along
under the name the adapter derives:

```r
JBrowseR(
  assembly = "hg38",
  tracks = list(list(uri = "peaks.bed.gz", name = "Peaks")),
  local_files = "~/analysis/peaks.bed.gz" # picks up peaks.bed.gz.tbi beside it
)
```

The bytes are base64-encoded into the document, so this is for a file on your own
machine rather than for a reference genome; `local_files` warns past ~50 MB, and
the [Hosting data](../creating-urls.html) vignette covers what to do above that.

## A statistical plot (GWAS Manhattan)

A track isn't only features and coverage — a display can plot data. Point a
[`GWASTrack`](https://jbrowse.org/jb2/docs/config/gwasadapter/) at genome-wide
summary statistics and give it a
[`LinearManhattanDisplay`](https://jbrowse.org/jb2/docs/config/linearmanhattandisplay/),
and the same linear view becomes a Manhattan plot: the `displays` block picks the
plot; no special widget needed. The adapter's `uri` shorthand finds the `.tbi`
index next to the data, and JBrowse fills in `displayId`, so the config stays
short.

```{r manhattan}
JBrowseR(
  assembly = "hg19",
  tracks = list(list(
    type = "GWASTrack",
    trackId = "gwas_track",
    name = "GWAS",
    adapter = list(
      type = "GWASAdapter",
      scoreColumn = "neg_log_pvalue",
      uri = "https://jbrowse.org/genomes/hg19/gwas/summary_stats.txt.gz"
    ),
    displays = list(list(type = "LinearManhattanDisplay", height = 250))
  )),
  location = "2"
)
```

For comparative genomics — synteny and dotplots across two assemblies — see the
[comparing genomes](synteny.html) article, which uses `JBrowseRApp()`.
