---
title: "Live embedded browsers"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Every browser on this page is a **live, interactive JBrowse 2** — pan, zoom, and
click features directly in the page. This article is part of the website only
(it is not a package vignette), so it can embed the full widget; the
[Introduction](../JBrowseR.html) uses screenshots to stay within CRAN's size
limits.

```{r setup}
library(JBrowseR)
```

## A hosted genome

Name a genome hub and you get the assembly, reference-name aliases, cytobands,
and gene-name search for free. Here `location = "BRCA1"` resolves through the
hub's search index.

```{r hub}
JBrowseR(
  "hg38",
  tracks = tracks(track(
    paste0(
      "https://s3.amazonaws.com/jbrowse.org/genomes/GRCh38/ncbi_refseq/",
      "GCA_000001405.15_GRCh38_full_analysis_set.refseq_annotation.sorted.gff.gz"
    ),
    name = "NCBI RefSeq Genes"
  )),
  location = "BRCA1"
)
```

## A custom genome from a FASTA URL

No hub needed for your own genome: pass a sequence-file URL straight to
`assembly`. The view builds the assembly from it (naming it after the file), and
a bare data-file URL in `tracks` is inferred into a track — here an alignments
track from a BAM.

```{r custom}
JBrowseR(
  "https://jbrowse.org/genomes/volvox/volvox.fa",
  tracks = list("https://jbrowse.org/genomes/volvox/volvox.bam"),
  location = "ctgA:1..5,000"
)
```

## An R result as a track

Turn a data frame of features into a track with `track_data_frame()` — no files,
no web server. A `score` column makes it quantitative.

```{r dataframe}
peaks <- data.frame(
  chrom = "17",
  start = seq(43000000, 43120000, by = 12000),
  end = seq(43000000, 43120000, by = 12000) + 4000,
  name = paste0("peak", seq_len(11)),
  score = round(runif(11, 5, 100))
)

JBrowseR(
  "hg38",
  tracks = list(track_data_frame(peaks, "R_peaks")),
  location = "17:43,000,000..43,125,000"
)
```

## A statistical plot (GWAS Manhattan)

A track isn't only features and coverage — a display can plot data. Point a
`GWASTrack` at genome-wide summary statistics and give it a
`LinearManhattanDisplay`, and the same linear view becomes a Manhattan plot:
the `displays` block picks the plot; no special widget needed.

```{r manhattan}
JBrowseR(
  "hg19",
  tracks = list(list(
    type = "GWASTrack",
    trackId = "gwas_track",
    name = "GWAS",
    assemblyNames = list("hg19"),
    adapter = list(
      type = "GWASAdapter",
      scoreColumn = "neg_log_pvalue",
      bedGzLocation = list(
        uri = "https://s3.amazonaws.com/jbrowse.org/genomes/hg19/gwas/summary_stats.txt.gz"
      ),
      index = list(location = list(
        uri = "https://s3.amazonaws.com/jbrowse.org/genomes/hg19/gwas/summary_stats.txt.gz.tbi"
      ))
    ),
    displays = list(list(
      type = "LinearManhattanDisplay",
      displayId = "gwas_track-manhattan",
      height = 250
    ))
  )),
  location = "2"
)
```

For comparative genomics — synteny and dotplots across two assemblies — see the
[comparing genomes](synteny.html) article, which uses `JBrowseRApp()`.
