---
title: "Comparing genomes: synteny and dotplots"
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup}
library(JBrowseR)
```

`JBrowseR()` shows a single linear view. For anything comparative — synteny,
dotplots, breakpoint-split — `JBrowseRApp()` drives the full JBrowse 2 app, and a
`views` list describes what to open. Each entry is a `list(type, init)` spec, the
same vocabulary JBrowse Web serializes into its
[`?session=spec-…` URLs](https://jbrowse.org/jb2/docs/urlparams/), built with
`linear_view()`, `synteny_view()`, or `dotplot_view()`. The fields each `init`
accepts come from the view's
[state-model docs](https://jbrowse.org/jb2/docs/models/) — see
[LinearSyntenyView](https://jbrowse.org/jb2/docs/models/linearsyntenyview/) — and
the PAF track is a
[`PAFAdapter`](https://jbrowse.org/jb2/docs/config/pafadapter/).

The browser below is **live** — drag either genome to pan; the ribbon between
them follows.

## Linear synteny

Two `volvox` assemblies (one carries a deletion) are tied together by a PAF
track and shown as a linear synteny view. `assembly()` handles the `.2bit` and
FASTA sequences; `synteny_track()` builds the PAF track that spans both.

```{r synteny}
base <- "https://jbrowse.org/code/jb2/main/test_data/volvox/"

JBrowseRApp(
  assemblies = list(
    assembly(paste0(base, "volvox.2bit"), name = "volvox"),
    assembly(paste0(base, "volvox_del.fa"), name = "volvox_del")
  ),
  tracks = list(
    synteny_track(paste0(base, "volvox_del.paf"), "volvox", "volvox_del",
      track_id = "volvox_del_paf")
  ),
  views = list(synteny_view(c("volvox", "volvox_del"), tracks = "volvox_del_paf"))
)
```

## Dotplot

The same assemblies and PAF render as a dotplot with one function call
different — `dotplot_view()` in place of `synteny_view()`:

```{r dotplot}
JBrowseRApp(
  assemblies = list(
    assembly(paste0(base, "volvox.2bit"), name = "volvox"),
    assembly(paste0(base, "volvox_del.fa"), name = "volvox_del")
  ),
  tracks = list(
    synteny_track(paste0(base, "volvox_del.paf"), "volvox", "volvox_del",
      track_id = "volvox_del_paf")
  ),
  views = list(dotplot_view(c("volvox", "volvox_del"), tracks = "volvox_del_paf"))
)
```
