---
title: "Comparing genomes: synteny and dotplots"
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup}
library(JBrowseR)
```

`JBrowseR()` shows a single linear view. For anything comparative — synteny,
dotplots, breakpoint-split — `JBrowseRApp()` drives the full JBrowse 2 app, and a
`views` list describes what to open. Each entry is a `list(type, init)` spec, the
same vocabulary JBrowse Web serializes into its
[`?session=spec-…` URLs](https://jbrowse.org/jb2/docs/urlparams/) — a plain
list, with no R builder in the way. The fields each `init` accepts come from the
view's
[state-model docs](https://jbrowse.org/jb2/docs/models/) — see
[LinearSyntenyView](https://jbrowse.org/jb2/docs/models/linearsyntenyview/) — and
the PAF track is a
[`PAFAdapter`](https://jbrowse.org/jb2/docs/config/pafadapter/).

The browser below is **live** — drag either genome to pan; the ribbon between
them follows.

## Linear synteny

Two `volvox` assemblies (one carries a deletion) are tied together by a PAF
track and shown as a linear synteny view. Each assembly is a `list(name=, uri=)`
that core expands; the PAF track spans both and so names them in order.

```{r synteny}
base <- "https://jbrowse.org/code/jb2/main/test_data/volvox/"

JBrowseRApp(
  assemblies = list(
    list(name = "volvox", uri = paste0(base, "volvox.2bit")),
    list(name = "volvox_del", uri = paste0(base, "volvox_del.fa"))
  ),
  tracks = list(
    list(
      type = "SyntenyTrack",
      trackId = "volvox_del_paf",
      name = "volvox_del.paf",
      assemblyNames = list("volvox", "volvox_del"),
      adapter = list(
        type = "PAFAdapter",
        targetAssembly = "volvox",
        queryAssembly = "volvox_del",
        uri = paste0(base, "volvox_del.paf")
      )
    )
  ),
  views = list(list(
    type = "LinearSyntenyView",
    init = list(
      views = list(list(assembly = "volvox"), list(assembly = "volvox_del")),
      tracks = list("volvox_del_paf")
    )
  ))
)
```

## Dotplot

The same assemblies and PAF render as a dotplot with one field different —
`"DotplotView"` in place of `"LinearSyntenyView"`:

```{r dotplot}
JBrowseRApp(
  assemblies = list(
    list(name = "volvox", uri = paste0(base, "volvox.2bit")),
    list(name = "volvox_del", uri = paste0(base, "volvox_del.fa"))
  ),
  tracks = list(
    list(
      type = "SyntenyTrack",
      trackId = "volvox_del_paf",
      name = "volvox_del.paf",
      assemblyNames = list("volvox", "volvox_del"),
      adapter = list(
        type = "PAFAdapter",
        targetAssembly = "volvox",
        queryAssembly = "volvox_del",
        uri = paste0(base, "volvox_del.paf")
      )
    )
  ),
  views = list(list(
    type = "DotplotView",
    init = list(
      views = list(list(assembly = "volvox"), list(assembly = "volvox_del")),
      tracks = list("volvox_del_paf")
    )
  ))
)
```
